### 2022-04-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@07bf7fcfd0bbd09f9ae2dcefa0c47eeaf7e27313 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@29ac99b98a2aab44e530d1a13ee42e5a12e4ae6c 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@8dd7d67ed0c1ccefb16231d48b0addedccf9764e 🚀

Bats are important reservoir hosts for different classes of
microorganisms (Chu 2008, Donaldson 2010, Li 2010), some of which can
threaten human health. Chiropterans emerged around 64 million years ago
and are one of the most diverse mammalian orders, with an estimated
richness of more than 12000 species, (Peixoto et al. 2018) and 14325
known species (Simmons and Cirranello 2020). They exhibit a broad
variety of habitat use, behaviour, and feeding strategies, resulting in
their playing an essential role in the delivery of several ecosystem
services tied to important ecosystem-derived benefits (Kasso and
Balakrishnan 2013). For example, over two-thirds of bats are know to be
either obligate or facultative insectivorous mammals, therefore playing
an important role in the regulation of insect pests that can affect
crops (Williams-Guillén et al. 2008, Voigt and Kingston 2016), and
vectors of diseases that put a risk on human health (Gonsalves et al.
2013a, b). Because bats are globally distributed and have a long
evolutionary history, phylogeographic and biogeographic approaches are
required to shed light on the extant distribution of coevolutionary
processes between bats and the pathogens they carry. Not all areas in
which bats, viruses, and human are co-occuring are facing a risk of
spillover towards human populations, and the areas in which this risk
exist may not be facing risks of the same nature and magnitude.

@@

\hypertarget{viral-phylogeography}{%
\subsection{Viral phylogeography}\label{viral-phylogeography}}

We used the following query to pull all betacoronavirus sequence data
from the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent,''
``mutant,'' ``GFP,'' and ``recombinant.'' We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
(Katoh and Standley 2013, Supplemental X) and a maximum likelihood tree
reconstructed in IQ-TREE v 1.6.12 (Nguyen et al.~2015) with ModelFinder
(Kalyaanamoorthy et al.~2017) ultrafast bootstrap approximation (Hoang
et al. 2018) and the following parameters (STEPH WILL ADD, Supplemental
X).

\hypertarget{viral-evolutionary-diversification}{%
\subsection{Viral evolutionary
diversification}\label{viral-evolutionary-diversification}}

@@

\hypertarget{outbreaks-data-geo-referencing}{%
\subsection{Outbreaks data
geo-referencing}\label{outbreaks-data-geo-referencing}}

\hypertarget{results}{%
\section{Results}\label{results}}

@@

Chiroptera are an hyperdiverse group, distributed in a large part if the
world, and are an important reservoir for different strains of
betacoronaviruses (Drexler et al. 2014); this has attracted attention to
areas where high diversity of bats can be an important issue for human
health (Calisher et al. 2006, Moratelli and Calisher 2015). Accordingly,
we collected the IUCN rangemaps for known hosts of betacoronaviruses, to
illustrate where hotspots of host diversity are. These results are
presented in Fig xx.a. As per our current knowledge of which bats are
hosts of betacoronaviruses, these hotspots are primarily South-East
Asia, parts of Europe, and to a lesser extent sub-saharan Africa. Even
the subset of chiroptera that are hosts of betacoronaviruses fits the
evolutionary timeline of the group. Chiropterans can be classified as
Microchiroptera and macrochiroptera, where macrochiroptera have an older
history from an evolutionary perspective compared to macrochiroptera
(Teeling et al. 2005, Springer 2013). South-East Asia has a high
diversity of bats (Kingston, 2010), and our results show that part of
that diversity includes betacoronavirus hosts. High density of hosts
sharing the same virus (albeit possibly different strains) calls into
question the evolution of the bat antiviral immune system and its
co-evolution with viruses, which may result in distinct immunological
responses in different area, as evidenced in other bat species (Banerjee
et al. 2020).

@@

Higher host diversity may not result in a higher viral diversity; for
this reason, we quantified and mapped the evolutionary distinctiveness
of betacoronaviruses, based on \ldots. Viral evolutionary
distinctiveness largely tracks host diversity, particularly in southern
China but oddly not throughout the rest of southeast Asia, perhaps
indicating that many distinctive viruses remain to be discovered in this
region (an idea that is unsurprising given the growing realization,
around the emergence of SARS-CoV-2, that a unique lineage of similar
viruses are widespread in bats but still mostly undescribed). The most
distinct betacoronaviruses are found in South America, a region with a
comparatively lower number of hosts; this suggests that the South
American bat-betacoronvirus complex has been more isolated, and is
probably undergoing a different co-evolutionary dynamic. Alternatively,
this distinctiveness hostpot may be a product of under-sampling:
South-America is one of the places where the fewest betacoronaviruses
have been discovered (Anthony et al. 2017), and adding more viruses
would bring the distinctiveness of known sequences down. Previous work
has suggested the Americas may be a hotspot of both undiscovered bat
viruses in general (Olival et al. 2017, Allen et al. 2017) and
coronavirus specifically (Anthony et al. 2017), though not necessarily
betacoronaviruses, and particularly not those in clades with notable
zoonotic potential.

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

Bats are found worldwide and are one of the most diverse groups among
mammals (Moratelli \& Calisher, 2015). Previous research (Anthony et
al., 2017; Mollentze \& Streicker, 2020) states that locally diverse bat
communities could maintain more viruses and hence, a higher probability
of having a pathogen that could represent a risk for human health. This
probability involves multiple factors, among which the relatedness of
hosts (which can make the jumps easier (Longdon et al., 2011; Mollentze
et al., 2020; Wolfe et al., 2007), and the overall tendency of hosts
within a locality to share viruses, which may limit viral diversity
because of within-host competition (Leeks et al., 2018; Sallinen et al.,
2020). Species richness, therefore, is not a sufficient measure of viral
risk. This is exemplified in our results, where both South America and
South-Eastern Asia have a high species richness of betacov hosts, but
only the latter region has a high risk. Specifically, because previous
studies propose that Asia is important when it comes to understanding
the evolutionary origin of various mammalian taxa (Beard C K, 1988).
Including bats (Yu et al., 2014), which could support the relationship
between evolutionary time and the development of an immune system with
characteristics that allow them to be better adapted to infection by
emerging viruses (Gorbunova et al., 2020; Irving et al., 2021) may be
related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013).

@@

Comparing scenarios of high viral exchange vs low viral exchange, open
the discussion to consider if the best scenario is where viruses easily
adapted to multiple hosts but with low virulence or easily ignored by
the immune system of the host, or where we have viruses specialized to a
specific host, but highly virulent when invade a new host. Accordingly,
the understanding of viral-host interactions from a taxonomic and
phylogenetic contributes to improving zoonoses surveillance programs.

>>>>

Bats are important reservoir hosts for different classes of
microorganisms (Chu 2008, Donaldson 2010, Li 2010), some of which can
threaten human health. Chiropterans emerged around 64 million years ago
and are one of the most diverse mammalian orders, with an estimated
richness of more than 14000 species (Peixoto et al. 2018, Simmons and
Cirranello 2020). They exhibit a broad variety of habitat use,
behaviour, and feeding strategies, resulting in their playing an
essential role in the delivery of several ecosystem services tied to
important ecosystem-derived benefits (Kasso and Balakrishnan 2013). For
example, over two-thirds of bats are know to be either obligate or
facultative insectivorous mammals, therefore playing an important role
in the regulation of insect pests that can affect crops
(Williams-Guillén et al. 2008, Voigt and Kingston 2016), and vectors of
diseases that put a risk on human health (Gonsalves et al. 2013a, b).
Because bats are globally distributed and have a long evolutionary
history, phylogeographic and biogeographic approaches are required to
shed light on the extant distribution of coevolutionary processes
between bats and the pathogens they carry. Not all areas in which bats,
viruses, and human are co-occuring are facing a risk of spillover
towards human populations, and the areas in which this risk exist may
not be facing risks of the same nature and magnitude.

@@

\hypertarget{viral-phylogeography-and-evolutionary-diversification}{%
\subsection{Viral phylogeography and evolutionary
diversification}\label{viral-phylogeography-and-evolutionary-diversification}}

We used the following query to pull all betacoronavirus sequence data
from the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent,''
``mutant,'' ``GFP,'' and ``recombinant.'' We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
(\textbf{Katoh and Standley 2013}, parameters in text?) and a maximum
likelihood tree reconstructed in IQ-TREE v 1.6.12 (Nguyen et al.~2015)
with ModelFinder (Kalyaanamoorthy et al.~2017) ultrafast bootstrap
approximation (Hoang et al. 2018) and the following parameters
(\textbf{STEPH WILL ADD}, parameters in text?).

@@

\hypertarget{results-and-discussion}{%
\section{Results and discussion}\label{results-and-discussion}}

@@

Bats are found worldwide and are one of the most diverse groups among
mammals (Moratelli \& Calisher, 2015), and one of the main animal
reservoir for different strains of betacoronaviruses (Drexler et al.
2014). This has attracted attention to areas where high diversity of
bats, and therefore presumably high diversity of betacoronaviruses, can
be an important issue for human health (Calisher et al. 2006, Moratelli
and Calisher 2015). Accordingly, we collected the IUCN rangemaps for
known hosts of betacoronaviruses, to illustrate where hotspots of host
diversity are. These results are presented in Fig xx.a. As per our
current knowledge of which bats are hosts of betacoronaviruses, these
hotspots are primarily South-East Asia, parts of Europe, and to a lesser
extent sub-saharan Africa. Even the subset of chiroptera that are hosts
of betacoronaviruses fits the evolutionary timeline of the group.
Chiropterans can be classified as Microchiroptera and macrochiroptera,
where macrochiroptera have an older history from an evolutionary
perspective compared to macrochiroptera (Teeling et al. 2005, Springer
2013).

South-East Asia has a high diversity of bats (\textbf{Kingston, 2010}),
and our results show that part of that diversity includes
betacoronavirus hosts. High density of hosts sharing the same virus
(albeit possibly different strains) calls into question the evolution of
the bat antiviral immune system and its co-evolution with viruses, which
may result in distinct immunological responses in different areas, as
evidenced in other bat species (Banerjee et al. 2020). Immune
characteristics that allow bats to be better adapted to infection by
emerging viruses (Gorbunova et al., 2020; Irving et al., 2021) may be
related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013).
Considering whether viruses easily adapted to multiple hosts have lower
virulence on these hosts, or lower ability to jump to hosts with
different immune characteristics, should yield valuable additional
predictors for the total risk of spillover. Previous research
(\textbf{Anthony et al., 2017; Mollentze \& Streicker, 2020}) states
that locally diverse bat communities could maintain more viruses and
hence, a higher probability of having a pathogen that could represent a
risk for human health; locally diverse, virus-rich bats communities
could represent an increased risk of spillover under climate change
(\textbf{Ice ice berg berg}). This probability involves multiple
factors, among which the relatedness of hosts (which can make the jumps
easier (\textbf{Longdon et al., 2011; Mollentze et al., 2020; Wolfe et
al., 2007}), and the overall tendency of hosts within a locality to
share viruses, which may limit viral diversity because of within-host
competition (Leeks et al., 2018; Sallinen et al., 2020). All things
considered, the richness of known betacoronaviruses hosts is not a
sufficient predictor of spillover risk.

@@

Higher host diversity may not result in a higher viral diversity, for
example if all hosts share the same viruses, or share closely
evolutionarily related strains. For this reason, we quantified and
mapped the evolutionary distinctiveness of betacoronaviruses, based on
their position in a molecular phylogeny. Viral evolutionary
distinctiveness largely tracks host diversity, particularly in southern
China but, oddly, not throughout the rest of southeast Asia. This
indicates, perhaps, that many distinctive viruses remain to be
discovered in this region (an idea that is unsurprising given the
growing realization, around the emergence of SARS-CoV-2, that a unique
lineage of similar viruses are widespread in bats but still mostly
undescribed). The most distinct betacoronaviruses are found in South
America, a region with a comparatively lower number of hosts; this
suggests that the South American bat-betacoronvirus complex has been
more isolated, and is probably undergoing a different co-evolutionary
dynamic. Alternatively, this distinctiveness hostpot may be a product of
under-sampling: South-America is one of the places where the fewest
betacoronaviruses have been discovered (Anthony et al. 2017), and adding
more viruses would bring the distinctiveness of known sequences down.
Previous work has suggested the Americas may be a hotspot of both
undiscovered bat viruses in general (Olival et al. 2017, Allen et al.
2017) and coronavirus specifically (Anthony et al. 2017), though not
necessarily betacoronaviruses, and particularly not those in clades with
notable zoonotic potential.

@@

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

@@

Species richness, therefore, is not a sufficient measure of viral risk.
This is exemplified in our results, where both South America and
South-Eastern Asia have a high species richness of betacov hosts, but
only the latter region has a high risk. Specifically, because previous
studies propose that Asia is important when it comes to understanding
the evolutionary origin of various mammalian taxa (Beard C K, 1988).

@@
